---
title: "Run_preps_cleaned"
author: "Annemarie Eckes-Shephard"
date: "2025-06-11"
output: html_document
---

```{r setup, include=FALSE}

library(ggplot2)
library(sf)

```


Needs more cleaning, but these are the essential input file preparation and exploration steps I took for ForestPaths sims.

## 21.01.2025 creation of gridlist file
There must be some missing gridcells in the simulation, which fails in places:
I gapfilled some of Susanna's grids to make compatible with the LU file from Martin.

```{r gridlist_checks_and_create_gridlist}
# Load the maps package
library(maps)
# Load necessary libraries
library(dplyr)
# Set up the plotting layout for 4 maps (2 rows, 2 columns)
par(mfrow = c(2, 2), mar = c(2, 2, 2, 2))  # Reduced margins

############################################################
# Combine broadleaf and conifer lat lons
Forest_harv_coords_BL <- read.table(file = "../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/intensity_mean_Broadleaf.txt",header=TRUE)
coords_BL <- unique(Forest_harv_coords_BL[,c("lon","lat")],)
Forest_harv_coords_CO <- read.table(file = "../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/intensity_mean_Conifer.txt",header=TRUE)
coords_CO<- unique(Forest_harv_coords_CO[,c("lon","lat")],)

tmp <- rbind(coords_CO,coords_BL)
Prob_harvest_latlon<- unique(tmp)
names(Prob_harvest_latlon) <- c("Lon","Lat")

# Set up the plot
map(
    "world",                      # Map data
    xlim = c(-25, 35),            # Longitude range
    ylim = c(35, 70),             # Latitude range
    col = "gray80",               # Color of land
    fill = TRUE,                  # Fill the polygons
    border = "gray50",            # Border color of countries
    bg = "lightblue",             # Background color (ocean)
    lwd = 0.5                    # Line width for borders
)

points(Prob_harvest_latlon$Lon,Prob_harvest_latlon$Lat, col=1,pch=16,cex=0.6)
mtext(side=3,"probabilistic harvest")


############################################################
#landuse change
#echo "Lon Lat" > latlon_lu.txt
#awk '$3 <= 2009 { print $1, $2 }' lucrop_hildaPucherMircaGenusmap.txt  > latlon_lu.txt
#awk '$3 <= 2009 { print $1, $2 }' luforest_hildaPucherMircaGenusmap.txt  >> latlon_lu.txt
lucrop <- unique(read.table("../landuse_change_forest_age/latlon_lu.txt",header = TRUE)[,c("Lon","Lat")])

# check against other landuse files:
luc <- unique(read.table("../landuse_change_forest_age/lu_inputfiles_noNF/net_lu_HildaPucherMircaEurope.txt",header = TRUE)[,c("Lon","Lat")])

# Set up the plot
map(
    "world",                      # Map data
    xlim = c(-25, 35),            # Longitude range
    ylim = c(35, 70),             # Latitude range
    col = "gray80",               # Color of land
    fill = TRUE,                  # Fill the polygons
    border = "gray50",            # Border color of countries
    bg = "lightblue",             # Background color (ocean)
    lwd = 0.5                    # Line width for borders
)

mtext(side=3,"landuse change")
points(lucrop$Lon,lucrop$Lat,col="purple",pch=16,cex=0.4)
points(luc$Lon,luc$Lat,col="red",pch=16,cex=0.4) ## check against other landuse input files:

#sanity-check they are indeed covering the same lat lon extent:
# Find combinations in lucrop but not in luc
in_lucrop_not_in_luc <- setdiff(luc, lucrop)

# Find combinations in luc but not in lucrop
in_luc_not_in_lucrop <- setdiff(lucrop, luc)

#both named listes with no content. So there is a total overlap between the two datasets. that's good
#in_luc_not_in_lucrop 
#in_lucrop_not_in_luc

############################################################
#current gridlist in use: 
# Jun: nb this is no longer the "current" gridlist. but is expanded to accommodate the whole Prob harvest, Lu extent.
gridlist <- read.table("../european_applications/testrun/gridlist_fix_europe_rand_plus_norway_26gc.txt")
names(gridlist) <- c("Lon","Lat")
# Set up the plot
map(
    "world",                      # Map data
    xlim = c(-25, 35),            # Longitude range
    ylim = c(35, 70),             # Latitude range
    col = "gray80",               # Color of land
    fill = TRUE,                  # Fill the polygons
    border = "gray50",            # Border color of countries
    bg = "lightblue",             # Background color (ocean)
    lwd = 0.5                    # Line width for borders
)

mtext(side=3,"gridlist")
points(gridlist$Lon,gridlist$Lat,col="blue",pch=10,cex=0.4)
############################################################


############################################################
#disturbance input 
#on Simba in the disturbance input folder:
#echo "Lon Lat" > dist_coords.txt
#awk 'NR >1 { print $1, $2 }'  disturbance_0.5deg_harvest_use.txt >> dist_coords.txt
#awk 'NR >1 { print $1, $2 }'  disturbance_0.5deg_wind_use.txt >> dist_coords.txt
#awk 'NR >1 { print $1, $2 }'  disturbance_0.5deg_barkbeetle_use.txt >> dist_coords.txt
#awk 'NR > 1 { print $1, $2 }' disturbance_0.5deg_fire_use.txt >> dist_coords.txt
#disturbance coords
distcoords <- unique(read.table("../disturbance/dist_coords.txt",header = TRUE)[,c("Lon","Lat")])

# Set up the plot
map(
    "world",                      # Map data
    xlim = c(-25, 35),            # Longitude range
    ylim = c(35, 70),             # Latitude range
    col = "gray80",               # Color of land
    fill = TRUE,                  # Fill the polygons
    border = "gray50",            # Border color of countries
    bg = "lightblue",             # Background color (ocean)
    lwd = 0.5                    # Line width for borders
)

mtext(site=3,"disturbance")
points(distcoords$Lon,distcoords$Lat,col="orange",pch=10,cex=0.4)
############################################################

mtext(outer=TRUE, side=3, "Inpute file- coordinate extents", line=-2)



#crossover between prob_harvest and landuse:
 
# Set up the plot
map(
    "world",                      # Map data
    xlim = c(-25, 35),            # Longitude range
    ylim = c(35, 70),             # Latitude range
    col = "gray80",               # Color of land
    fill = TRUE,                  # Fill the polygons
    border = "gray50",            # Border color of countries
    bg = "lightblue",             # Background color (ocean)
    lwd = 0.5                    # Line width for borders
)
 
points(Prob_harvest_latlon$Lon,Prob_harvest_latlon$Lat, col=1,pch=16,cex=0.6)
points(luc$Lon,luc$Lat,col="red",pch=16,cex=0.4) ## check against other landuse input files:


shared_by_both <- inner_join(Prob_harvest_latlon, luc, by = c("Lat", "Lon"))
points(shared_by_both$Lon,shared_by_both$Lat,col="magenta",pch=16,cex=0.3)



# Step 1: Get rows unique to Prob_harvest_latlon (not in luc)
unique_to_prob_harvest <- anti_join(Prob_harvest_latlon, luc, by = c("Lat", "Lon"))
points(unique_to_prob_harvest$Lon,unique_to_prob_harvest$Lat,col="green",pch=16,cex=0.6)

# Step 2: Get rows unique to luc (not in Prob_harvest_latlon)
unique_to_luc <- anti_join(luc, Prob_harvest_latlon, by = c("Lat", "Lon"))
points(unique_to_luc$Lon,unique_to_luc$Lat,col="yellow",pch=16,cex=0.6)

# Step 3: Combine the unique rows from both data frames (union)
unique_latlon <- bind_rows(unique_to_prob_harvest, unique_to_luc)

#sanity check that it covers everything:
#points(unique_latlon$Lon,unique_latlon$Lat, col="grey", cex=0.5)

plot.new()
legend(legend=c("crossover","unique_to_prob_h","unique_to_luc"),fill=c("magenta","green","yellow"),"topleft")

##create gridlist that contains the landuse-change map as basis: 
write.table(file="../european_applications/testrun/gridlist_final.txt",luc, col.names =FALSE,row.names=FALSE)

#some points that fail (March 2025:)
# points(20.2500, 68.7500)
# points(15.7500, 66.2500)
# points(17.7500, 67.7500)
# points(-1.250, 49.250)
# points(20.75, 45.75)
#lat= 68.75
#lon= 20.25
#exists <- any(Prob_harvest_latlon$Lon == lon & Prob_harvest_latlon$Lat == lat)
#print(exists)
#exists <- any(luc$Lon ==lon & luc$Lat == lat)
#print(exists)
#exists <- any(unique_to_prob_harvest$Lon == lon & unique_to_prob_harvest$Lat == lat)
#print(exists)
#exists <- any(unique_to_luc$Lon == lon & unique_to_luc$Lat == lat)
#print(exists)

```


To create a gridlist that contains all complete latitude-longitude combinations of the landuse change-map, 
we have to gap-fill probabilistic harvest for 87 locations on the map. To identify the datapoints with which to gap fill, we use the Haversine distance formula. This formula computes the distance between two points on the Earth's surface given their latitude and longitude (black dots in the above plot are the outcome)

```{r gap_fill_prob_harvest}
library(geosphere)

#visualise problem: lat lons don't fully overlap:
map(
    "world",                      # Map data
    xlim = c(-25, 35),            # Longitude range
    ylim = c(35, 70),             # Latitude range
    col = "gray80",               # Color of land
    fill = TRUE,                  # Fill the polygons
    border = "gray50",            # Border color of countries
    bg = "lightblue",             # Background color (ocean)
    lwd = 0.5                    # Line width for borders
)
 
points(Prob_harvest_latlon$Lon,Prob_harvest_latlon$Lat, col=1,pch=16,cex=0.6)
points(luc$Lon,luc$Lat,col="red",pch=16,cex=0.4) ## check against other landuse input files:

# "wrong: full join not useful here"
# shared_by_both <- full_join(Prob_harvest_latlon, luc, by = c("Lat", "Lon"))
# points(shared_by_both$Lon,shared_by_both$Lat,col="magenta",pch=16,cex=0.3)


# Step 1: Get rows unique to Prob_harvest_latlon (not in luc)
unique_to_prob_harvest <- anti_join(Prob_harvest_latlon, luc, by = c("Lat", "Lon"))
points(unique_to_prob_harvest$Lon,unique_to_prob_harvest$Lat,col="green",pch=16,cex=0.6)

# Step 2: Get rows unique to luc (not in Prob_harvest_latlon)
unique_to_luc <- anti_join(luc, Prob_harvest_latlon, by = c("Lat", "Lon"))
points(unique_to_luc$Lon,unique_to_luc$Lat,col="yellow",pch=16,cex=0.6)

# we need to consider the "yellow dots" in the probabilistic harvest dataframe, too.

####################################################################################################################
#find what forest type and harvest intensity is in nearest neighbour cells:
#forest type:

Forest_harv_coords_BL <- read.table(file = "../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/intensity_mean_Broadleaf.txt",header=TRUE)
coords_BL <- unique(Forest_harv_coords_BL[,c("lon","lat")],)

Forest_harv_coords_CO <- read.table(file = "../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/intensity_mean_Conifer.txt",header=TRUE)
coords_CO<- unique(Forest_harv_coords_CO[,c("lon","lat")],)

# Create a function to find closest coordinates - distVincentySphere
find_closest_coordinates <- function(unique_to_luc, Forest_harv_coords_BL) {
  # Initialize a dataframe to store closest coordinates
  closest_coords <- data.frame(luc_lat = numeric(), luc_lon = numeric(),
                               lat = numeric(), lon = numeric(), dist = numeric())
  
  # Iterate through each row in unique_to_luc
  for (i in 1:nrow(unique_to_luc)) {
    # Get lat and lon of the current point in unique_to_luc
    luc_lat <- unique_to_luc$lat[i]
    luc_lon <- unique_to_luc$lon[i]
    
    # Calculate distances from the current point to all points in Forest_harv_coords_BL
    distances <- distVincentySphere(c(luc_lon, luc_lat), cbind(Forest_harv_coords_BL$lon, Forest_harv_coords_BL$lat))
    
    # Find the index of the closest point
    closest_index <- which.min(distances)
    
    # Store the closest coordinates, distance, and the original lat-lon from unique_to_luc
    closest_coords <- rbind(closest_coords, data.frame(luc_lat = luc_lat,
                                                       luc_lon = luc_lon,
                                                       lat = Forest_harv_coords_BL$lat[closest_index],
                                                       lon = Forest_harv_coords_BL$lon[closest_index],
                                                       dist = distances[closest_index]))
  }
  
  return(closest_coords)
}

# create function that uses the filename as input and then updates the file with the missing (unique_to_luc) lat lon combinations:
gapfill_using_closest_coordinates <- function(filename,unique_to_luc,show_it_works =FALSE){
  
    #just in case, adjust names to lower case:
    names(unique_to_luc) <- c("lon","lat")
    
    input_data_to_be_gapfilled <-  read.table(file = filename,header=TRUE)
    # Call the function and get the closest coordinates
    closest_coords <- find_closest_coordinates(unique_to_luc, input_data_to_be_gapfilled)
    
    # View the result
    if(show_it_works){
      points(closest_coords$lon,closest_coords$lat,col="black",cex=0.6,pch=16)
      legend(legend=c("crossover","unique_to_prob_h","unique_to_luc","nearest gridcell"),fill=c("magenta","green","yellow",1),"topleft",cex=0.7)
    }
    
    
    # merged_df contains after this the new lat lon combinations that were originally missing in Forest_harv_coords. Their data comes from the  lat lons that are the closest to the missing lat lons.
    merged_df <- merge(input_data_to_be_gapfilled,closest_coords, by = c("lat", "lon"))
    
    #remove the "old" lat and lon and rename the luc_coords to be the new lat lons. we then have duplicate content in all other columns, except for lat and lon for these closest_coords ( have gapfilled):
    merged_df$lat <- NULL
    merged_df$lon <- NULL
    merged_df$dist <- NULL
    # Rename luc_lat to lat and luc_lon to lon
    colnames(merged_df)[colnames(merged_df) == "luc_lat"] <- "lat"
    colnames(merged_df)[colnames(merged_df) == "luc_lon"] <- "lon"
    
    #now append to the Prob_harvest dataframe:
    complete_grid <- rbind(input_data_to_be_gapfilled,merged_df)
    
    if(show_it_works){
      points(complete_grid$lon,complete_grid$lat,cex=0.4,col="grey",pch=16)
    }
    
    # sanity check:
     #different_rows <- sum(!apply(complete_grid[1:3075,] == input_data_to_be_gapfilled, 1, all, na.rm = TRUE))
    #different_rows

   # Replace 0 values with "0.0000000"
    complete_grid_final <- data.frame(lapply(complete_grid, function(x) {
      if (is.numeric(x)) {
        # Format values: replace 0 with "0.0000000", keep others as they are
        ifelse(x == 0, format(0, nsmall = 7), x)
      } else {
        x
      }
    }))

    # write to file:
    write.table(complete_grid_final, file= paste0(filename,"_luc_coords_added"),row.names = FALSE,col.names=TRUE,quote=FALSE)

}


files <- c(
"../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/intensity_mean_Conifer.txt",
"../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/intensity_mean_Broadleaf.txt",
"../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/intensity_sd_Broadleaf.txt", 
"../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/intensity_sd_Conifer.txt",
"../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/p_clearcut_Broadleaf.txt",
"../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/p_clearcut_Conifer.txt",
"../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/p_harvest_Broadleaf.txt",
"../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/p_harvest_Conifer.txt",  
"../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/size_target_Broadleaf.txt",
"../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/size_target_Conifer.txt"     
)

for(filename in files){
  gapfill_using_closest_coordinates(filename,unique_to_luc)
}

# santiy_check:
ins <- read.table("../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/intensity_mean_Broadleaf.txt_luc_coords_added",head=TRUE)
points(ins$lon,ins$lat,col="dark green",pch=16,cex=0.3)
# it works! 
# [Q] is it a problem that Probharvest has more gridcells that gridlist? 
# this is now adressed anyways, by harmonising the gridlist



```

# landcover explorations

To get an understanding of what the landcover and genera input data looks like: 
What about the stand types and when do Wanda's genera come in?

```{r landuse_genus_area_change_explorations}

insss <- read.table("../landuse_change_forest_age/lu_inputfiles_noNF/luforest_HildaPucherMircaEurope.txt",header = TRUE)
insss_sub <- insss[which(insss$Lon==23.75 & insss$Lat ==35.25),]

# Pivoting from wide to long format
insss_long <- insss_sub %>%
  tidyr::pivot_longer(cols = c(Larix, Picea, Pinus, Fagus, Quercus, otherNL, otherBL, NatForest),
               names_to = "Forest_Cover", 
               values_to = "Fraction")
# Stacked area chart with smoothing
insss_long %>% 
  ggplot(aes(Year, Fraction, fill = Forest_Cover,  color = Forest_Cover)) +
  geom_area() 

#trying a couple more, to see whether this trend persists:
insss_sub <- insss[which(insss$Lon==5.75 & insss$Lat ==43.25),]

# Pivoting from wide to long format
insss_long <- insss_sub %>%
  tidyr::pivot_longer(cols = c(Larix, Picea, Pinus, Fagus, Quercus, otherNL, otherBL, NatForest),
               names_to = "Forest_Cover", 
               values_to = "Fraction")

# Stacked area chart with smoothing
insss_long %>% 
  ggplot(aes(Year, Fraction, fill = Forest_Cover,  color = Forest_Cover)) +
  geom_area() 

insss_sub <- insss[which(insss$Lon==31.75 & insss$Lat ==62.75),]

# Pivoting from wide to long format
insss_long <- insss_sub %>%
  tidyr::pivot_longer(cols = c(Larix, Picea, Pinus, Fagus, Quercus, otherNL, otherBL, NatForest),
               names_to = "Forest_Cover", 
               values_to = "Fraction")
# Stacked area chart with smoothing
insss_long %>% 
  ggplot(aes(Year, Fraction, fill = Forest_Cover,  color = Forest_Cover)) +
  geom_area() 


insss <- read.table("../landuse_change_forest_age/lu_inputfiles_noNF/net_lu_HildaPucherMircaEurope.txt",header = TRUE)
insss_sub <- insss[which(insss$Lon==5.75 & insss$Lat ==43.25),]

# Pivoting from wide to long format
insss_long <- insss_sub %>%
  tidyr::pivot_longer(cols = c(URBAN, CROPLAND, PASTURE, FOREST, NATURAL, BARREN),
               names_to = "Landusecover", 
               values_to = "Fraction")
# Stacked area chart with smoothing
insss_long %>% 
  ggplot(aes(Year, Fraction, fill = Landusecover,  color = Landusecover)) +
  geom_area() 




# if you need to read this in again:

insss_sub <- insss[which(insss$Year==2100),]
insss_sub_sf <- st_as_sf(insss_sub, coords = c('Lon', 'Lat'))
ggplot(insss_sub_sf) + 
  #geom_sf(aes(fill=NATURAL)) + scale_fill_viridis_c(option = "plasma", limits = c(0, 0.3)) 
  geom_tile(data = insss_sub, 
             aes(x = Lon, y = Lat, fill = NATURAL)) +  # Plot 'NATURAL' fraction as color
     scale_fill_viridis_c(option = "plasma", limits = c(0, 0.3)) +  # Adjust scale to 0-0.3
     coord_fixed() +  # Maintain correct aspect ratio
     labs(title = "Natural Land Fraction in 2100", 
          x = "Longitude", y = "Latitude", fill = "Fraction") +
     theme_minimal()

```

```{r plot_fractional_cover_genera}

ss <- read.table("../landuse_change_forest_age/lu_inputfiles_noNF2/net_lu_HildaPucherMircaEurope.txt",header = TRUE)
#lc 3 st fraction next year sum is zero. Check input data

ss_sub <- ss[which(ss$Year==2010),]
ss_sub <- ss[which(ss$Lon== 17.75 & ss$Lat==67.75),]
plot(ss_sub$URBAN)

ss <- read.table("/Users/annemarie/Documents/1_TreeMort/2_Analysis/3_analysis_ForestPaths_Exploratory_runs/landuse_change_forest_age/lu_inputfiles_noNF3/luforest_HildaPucherMircaEurope.txt",header = TRUE)
#-1.25,49.25
ss_sub <- ss[which(ss$Year>=1900 & ss$Year <= 1905 & ss$Lon== -1.25 & ss$Lat==49.25),]

ss_sub <- ss[which(ss$Year==2014),]


library(ggplot2)
library(dplyr)
library(tidyr)

# Reshape data to long format
ss_long <- ss_sub %>%
  pivot_longer(cols = Larix:otherBL, names_to = "Genus", values_to = "Fractional_Cover")

# Plot maps
ggplot(ss_long, aes(x = Lon, y = Lat, fill = Fractional_Cover)) +
  geom_tile() +
  scale_fill_viridis_c(option = "plasma", name = "Fractional Cover") + # Color scale
  facet_wrap(~Genus, ncol = 4) + # Arrange maps in a grid
  coord_fixed() + # Ensure correct aspect ratio
  labs(title = "Fractional Cover of Genera", x = "Longitude", y = "Latitude") +
  theme_minimal()

```



-------------------------------------------------------------------------------

# Management options

## Lighten and intensify thinning 

"Decrease intensity of all commercial thinnings by 25%, while holding rotation length fixed"

```{r process_and_visualise_probharv_intensity_changes}

#function to reduce all thinning intensity values by 25%:
reduce_by_25 <- function(df) {
  df[-c(1, 2)] <- lapply(df[-c(1, 2)], function(x) ifelse(x != 0, x * 0.75, 0))
  return(df)
}

#function to increase all thinning intensity values by 25%:
increase_by_25 <- function(df) {
  df[-c(1, 2)] <- lapply(df[-c(1, 2)], function(x) ifelse(x != 0, x * 1.25, 0))
  return(df)
}


# Define folder path
#this folder contains Susanna's Probabilistic harvest routine already distributed across the latlon combinations.
folder_path <- "../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/"
folder_path <- "../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/"

# List all files that start with "intensity_" and end with "_luc_coords_added"
file_list <- list.files(folder_path, pattern = "^intensity_.*_luc_coords_added$", full.names = TRUE)

# Function to process files
# writes directly to output file with updated name _red or _inc
process_file <- function(file) {
  # Read in the file
  df <- read.table(file, header = TRUE)

  # Apply transformations
  df_red <- reduce_by_25(df)
  df_inc <- increase_by_25(df)

  # Generate new filenames
  file_red <- paste0(file, "_red")
  file_inc <- paste0(file, "_inc")

  # Save the transformed data
  write.table(df_red, file = file_red, col.names = TRUE, row.names = FALSE,quote = FALSE)
  write.table(df_inc, file = file_inc, col.names = TRUE, row.names = FALSE,quote = FALSE)

  message("Processed: ", file)
}

# Apply the function to each file in the list
lapply(file_list, process_file)

########################
# sanity-check:
check_function_sd <- read.table(paste0(folder_path,"intensity_sd_Broadleaf.txt_luc_coords_added"),header = TRUE)
check_function_inc <- read.table(paste0(folder_path,"intensity_sd_Broadleaf.txt_luc_coords_added_inc"),header = TRUE)
check_function_red <- read.table(paste0(folder_path,"intensity_sd_Broadleaf.txt_luc_coords_added_red"),header = TRUE)

plot("n", xlim=c(-0.5,1), ylim=c(0,10),main="thinning\nintensity-occurrence\n across Europe", ylab="density", xlab="standard_deviation")
lines(density(check_function_sd$D25_BA0), col="black",lwd=2)
lines(density(check_function_inc$D25_BA0), col="red")
lines(density(check_function_red$D25_BA0), col="blue")
plot.new()
legend(legend=c("default","increased thinning", "decreased thinning"),fill=c(1,"red","blue"),"topright",cex=0.7)
```



## shorten and lengthen rotation periods

"Decrease or increase in rotation length, approximated by 25% increase or decrease of target diameter"

Right now I am increasing or decreasing the probability of harvest for a given diameter, that is different, isn't it? No, it's not ( see notes in chunk harvets_by_size_testing)


```{r process_and_visualise_probharv_size_target_changes}

#function to reduce all thinning intensity values by 25%:
reduce_by_25 <- function(df) {
  df[-c(1, 2)] <- lapply(df[-c(1, 2)], function(x) ifelse(x != 0, x * 0.75, 0))
  return(df)
}

#function to increase all thinning intensity values by 25%:
increase_by_25 <- function(df) {
  df[-c(1, 2)] <- lapply(df[-c(1, 2)], function(x) ifelse(x != 0, x * 1.25, 0))
  return(df)
}


# Define folder path
#this folder contains Susanna's Probabilistic harvest routine already distributed across the latlon combinations.
folder_path <- "../Probabilistic_harvest/lpjg_inputfiles_extrapolate_fmapOnly6_041224/"

# List all files that start with "intensity_" and end with "_luc_coords_added"
file_list <- list.files(folder_path, pattern = "^p_harvest.*_luc_coords_added$", full.names = TRUE)

# Function to process files
# writes directly to output file with updated name _red or _inc
process_file <- function(filename) {
  # Read in the file
  df <- read.table(filename, header = TRUE)

  # Apply transformations
  df_red <- reduce_by_25(df)
  df_inc <- increase_by_25(df)

  # Generate new filenames
  file_red <- paste0(filename, "_red")
  file_inc <- paste0(filename, "_inc")

  # Save the transformed data
  write.table(df_red, file = file_red, col.names = TRUE, row.names = FALSE,quote = FALSE)
  write.table(df_inc, file = file_inc, col.names = TRUE, row.names = FALSE,quote = FALSE)

  message("Processed: ", filename)
}

# Apply the function to each file in the list
lapply(file_list, process_file)

########################
# sanity-check:
check_function_sd  <- read.table(paste0(folder_path,"p_harvest_Broadleaf.txt_luc_coords_added"),header = TRUE)
check_function_inc <- read.table(paste0(folder_path,"p_harvest_Broadleaf.txt_luc_coords_added_inc"),header = TRUE)
check_function_red <- read.table(paste0(folder_path,"p_harvest_Broadleaf.txt_luc_coords_added_red"),header = TRUE)


plot("n", xlim=c(-0.05,0.1), ylim=c(0,50),main="harvest thresholds across Europe", ylab="density", xlab="standard_deviation")
lines(density(check_function_sd$D25_BA0), col="black",lwd=2)
lines(density(check_function_inc$D25_BA0), col="red")
lines(density(check_function_red$D25_BA0), col="blue")

legend(legend=c("default ","increased probability", "decreased probability"),fill=c(1,"red","blue"),"topright",cex=0.6)

```


# Gridlist- country mapping.
To avoid confusions when it comes to distribution of gridlists to European country names:


```{create_country_gridlist}

# Script to identify countries from latitude and longitude coordinates
# Required packages
if (!require("sf")) install.packages("sf")
if (!require("rnaturalearth")) install.packages("rnaturalearth")
if (!require("rnaturalearthdata")) install.packages("rnaturalearthdata")
if (!require("dplyr")) install.packages("dplyr")

library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(dplyr)

# Function to identify countries from lat-lon coordinates
add_country_names <- function(input_file) {
  # Read your data file (assuming CSV format with headers)
  # Adjust read.csv parameters based on your file format
  cat("Reading input file:", input_file, "\n")
  data <- read.table(input_file, stringsAsFactors = FALSE,head=FALSE)
  names(data) <- c("lon","lat")
  
  # Get world country data
  cat("Loading country boundary data...\n")
  world <- ne_countries(scale = "medium", returnclass = "sf")
  
  # Convert the data frame to an sf object
  cat("Processing coordinates...\n")
  points_sf <- st_as_sf(data, coords =  c("lon", "lat"), crs = 4326)
  
  # Perform spatial join to identify which country each point falls in
  cat("Matching coordinates to countries...\n")
  result <- st_join(points_sf, world %>% select(name, iso_a2, continent))
  
  # Convert back to a regular data frame
  result_df <- cbind(data, st_drop_geometry(result[, c("name", "iso_a2", "continent")]))
  
  # Rename columns to be more descriptive
  names(result_df)[names(result_df) == "name"] <- "country_name"
  names(result_df)[names(result_df) == "iso_a2"] <- "country_code"
  
  # Filter for European countries if needed
  # Uncomment the next line if you only want European countries
  # result_df <- result_df[result_df$continent == "Europe", ]
  
  # Write the result to a new CSV file
  output_file = paste0(input_file,"_country_names_added")
  cat("Writing output to:", output_file, "\n")
  write.table(result_df, output_file, row.names = FALSE)
  
  cat("Process completed. Added country information to", nrow(data), "coordinates.\n")
  cat("Points with identified countries:", sum(!is.na(result_df$country_name)), "\n")
  cat("Points without matching countries:", sum(is.na(result_df$country_name)), "\n")
  
  return(result_df)
}

########################################################################
##add gridlist files to which to add country names here:

input_files <- c("../european_applications/testrun/gridlist_final.txt","../european_applications/testrun/gridlist_stratified.txt")

# Run now:
for (input_file in input_files){
  add_country_names(input_file)
}

```



